---
title: "Netflix Showdown: Action vs. Romance"
subtitle: "Which genre has dominated Netflix releases over the past decade?"
author: "Data Science Project"
format:
  html:
    theme: darkly
    toc: true
execute:
  echo: false
  warning: false
  message: false
---

## The Showdown

Are we living in an era of high-octane explosions or heartfelt love stories? 
Let's explore Netflix's catalog to see whether Action or Romance films take the crown.

```{r}
#| label: setup
library(tidyverse)
library(plotly)
library(DT)

# Load the Kaggle Netflix dataset
netflix <- read_csv("netflix_titles.csv", show_col_types = FALSE)

# Filter for movies released between 2011 and 2020
films <- netflix |>
  filter(type == "Movie", release_year >= 2011, release_year <= 2020) |>
  mutate(
    genre = case_when(
      str_detect(listed_in, "Action & Adventure") & str_detect(listed_in, "Romantic Movies") ~ "Both",
      str_detect(listed_in, "Action & Adventure") ~ "Action",
      str_detect(listed_in, "Romantic Movies") ~ "Romance",
      TRUE ~ "Other"
    )
  ) |>
  filter(genre %in% c("Action", "Romance"))

counts_by_year <- films |>
  count(release_year, genre)
```

## Tale of the Tape: 10-Year Totals

```{r}
#| label: totals
total_action <- sum(films$genre == "Action")
total_romance <- sum(films$genre == "Romance")
```

Over the full 10-year window (2011–2020):

- 💥 **Action Films:** `r total_action`
- ❤️ **Romance Films:** `r total_romance`

Overall, **Action** edges out Romance in overall catalog volume! But how has this rivalry shifted year by year? Hover over the bars below to see annual releases:

```{r}
#| label: interactive-plot
p <- ggplot(counts_by_year, aes(
  x = factor(release_year), 
  y = n, 
  fill = genre,
  text = paste0("Year: ", release_year, "<br>Genre: ", genre, "<br>Films: ", n)
)) +
  geom_col(position = position_dodge(preserve = "single")) +
  scale_fill_manual(values = c("Action" = "#E50914", "Romance" = "#FF69B4")) +
  labs(
    title = "Annual Releases on Netflix (2011–2020)",
    x = "Release Year",
    y = "Number of Films Released",
    fill = "Genre"
  ) +
  theme_minimal()

ggplotly(p, tooltip = "text")
```

## The Cumulative Race Over Time

Did Action lead the entire time, or was it a neck-and-neck contest? Let's track the running cumulative total of releases year by year:

```{r}
#| label: cumulative-race
cumulative_data <- counts_by_year |>
  complete(release_year = 2011:2020, genre, fill = list(n = 0)) |>
  arrange(genre, release_year) |>
  group_by(genre) |>
  mutate(total_to_date = cumsum(n)) |>
  ungroup()

p_race <- ggplot(cumulative_data, aes(
  x = release_year, 
  y = total_to_date, 
  color = genre,
  group = genre,
  text = paste0("Year: ", release_year, "<br>Genre: ", genre, "<br>Total: ", total_to_date)
)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_color_manual(values = c("Action" = "#E50914", "Romance" = "#FF69B4")) +
  scale_x_continuous(breaks = seq(2011, 2020, by = 1)) +
  labs(
    title = "Cumulative Releases: Action vs Romance",
    x = "Year",
    y = "Cumulative Total Movies",
    color = "Genre"
  ) +
  theme_minimal()

ggplotly(p_race, tooltip = "text")
```

## Movie Explorer: Search & Filter

Want to see which movies are behind the numbers? Use the search bar below to find your favorite films, or filter by genre and year:

```{r}
#| label: movie-table
films |>
  select(Title = title, Genre = genre, Year = release_year, Rating = rating, Duration = duration, Description = description) |>
  datatable(
    filter = "top",
    options = list(
      pageLength = 8,
      autoWidth = TRUE,
      searchHighlight = TRUE
    ),
    rownames = FALSE,
    class = "cell-border stripe"
  )
```


