---
title: "Netflix Showdown: Action vs. Romance"
subtitle: "Which genre has dominated Netflix releases over the past decade?"
author: "Data Science Project"
format:
  html:
    theme: darkly
    toc: true
execute:
  echo: false
  warning: false
  message: false
---

## The Showdown

Are we living in an era of high-octane explosions or heartfelt love stories? 
Let's explore Netflix's catalog to see whether Action or Romance films take the crown.

```{r}
#| label: setup
library(tidyverse)
library(plotly)
library(DT)
library(jsonlite)

# Load the Kaggle Netflix dataset
netflix <- read_csv("netflix_titles.csv", show_col_types = FALSE)

# Filter for movies released between 2011 and 2020
films <- netflix |>
  filter(type == "Movie", release_year >= 2011, release_year <= 2020) |>
  mutate(
    genre = case_when(
      str_detect(listed_in, "Action & Adventure") & str_detect(listed_in, "Romantic Movies") ~ "Both",
      str_detect(listed_in, "Action & Adventure") ~ "Action",
      str_detect(listed_in, "Romantic Movies") ~ "Romance",
      TRUE ~ "Other"
    )
  ) |>
  filter(genre %in% c("Action", "Romance"))

counts_by_year <- films |>
  count(release_year, genre)
```

## Tale of the Tape: 10-Year Totals

```{r}
#| label: totals
total_action <- sum(films$genre == "Action")
total_romance <- sum(films$genre == "Romance")
```

Over the full 10-year window (2011–2020):

- 💥 **Action Films:** `r total_action`
- ❤️ **Romance Films:** `r total_romance`

Overall, **Action** edges out Romance in overall catalog volume! But how has this rivalry shifted year by year? Hover over the bars below to see annual releases:

```{r}
#| label: interactive-plot
p <- ggplot(counts_by_year, aes(
  x = factor(release_year), 
  y = n, 
  fill = genre,
  text = paste0("Year: ", release_year, "<br>Genre: ", genre, "<br>Films: ", n)
)) +
  geom_col(position = position_dodge(preserve = "single")) +
  scale_fill_manual(values = c("Action" = "#FF2A54", "Romance" = "#00E5FF")) +
  labs(
    title = "Annual Releases: Action vs Romance",
    x = "Release Year",
    y = "Films Released",
    fill = "Genre"
  ) +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#141414", color = NA),
    panel.background = element_rect(fill = "#141414", color = NA),
    text = element_text(color = "#FFFFFF"),
    axis.text = element_text(color = "#CCCCCC"),
    panel.grid.major = element_line(color = "#2A2A2A"),
    panel.grid.minor = element_blank()
  )

ggplotly(p, tooltip = "text") |>
  layout(
    paper_bgcolor = "#141414",
    plot_bgcolor = "#141414",
    font = list(color = "#FFFFFF")
  )
```

## The Animated Bar Race (2011–2020)

Press **Play** below to watch the race unfold year by year!

```{r}
#| label: animated-race
race_data <- counts_by_year |>
  complete(release_year = 2011:2020, genre, fill = list(n = 0)) |>
  arrange(genre, release_year) |>
  group_by(genre) |>
  mutate(total_to_date = cumsum(n)) |>
  ungroup()

plot_ly(
  race_data,
  x = ~total_to_date,
  y = ~genre,
  color = ~genre,
  colors = c("Action" = "#FF2A54", "Romance" = "#00E5FF"),
  frame = ~release_year,
  text = ~paste0(genre, ": ", total_to_date, " films"),
  hoverinfo = "text",
  type = "bar",
  orientation = "h"
) |>
  layout(
    paper_bgcolor = "#141414",
    plot_bgcolor = "#141414",
    xaxis = list(title = "Cumulative Total Films", range = c(0, 500), color = "#FFFFFF", gridcolor = "#2A2A2A"),
    yaxis = list(title = "", color = "#FFFFFF"),
    font = list(color = "#FFFFFF"),
    showlegend = FALSE
  ) |>
  animation_opts(frame = 800, transition = 350, redraw = FALSE) |>
  animation_button(x = 0, y = -0.2)
```

## 🍿 Mood Roulette: What Should You Stream Tonight?

Can't decide what to watch? Pick your vibe below and let the algorithm deal you a random movie from Netflix's 10-year collection!

```{r}
#| label: roulette-data
#| results: asis
films_subset <- films |>
  select(title, genre, release_year, rating, duration, description)

cat(paste0(
  "<script>const filmsData = ",
  toJSON(films_subset),
  ";</script>"
))
```

```{=html}
<div style="background: linear-gradient(145deg, #1f1f1f, #141414); border: 2px solid #2a2a2a; border-radius: 16px; padding: 28px; margin: 25px 0; text-align: center; box-shadow: 0 10px 30px rgba(0,0,0,0.6);">
  <p style="color: #e0e0e0; font-size: 1.1rem; margin-bottom: 20px;">
    Feeling adrenaline-fueled explosions, or heartstrings and butterflies?
  </p>
  
  <div style="display: flex; justify-content: center; gap: 16px; margin-bottom: 24px; flex-wrap: wrap;">
    <button id="btn-action" style="background: #FF2A54; color: #FFFFFF; border: none; padding: 12px 28px; border-radius: 50px; font-weight: 700; cursor: pointer; font-size: 1rem; box-shadow: 0 0 18px rgba(255, 42, 84, 0.45); transition: transform 0.15s ease;">
      💥 Spin Action Movie
    </button>
    <button id="btn-romance" style="background: #00E5FF; color: #0a0a0a; border: none; padding: 12px 28px; border-radius: 50px; font-weight: 700; cursor: pointer; font-size: 1rem; box-shadow: 0 0 18px rgba(0, 229, 255, 0.45); transition: transform 0.15s ease;">
      💖 Spin Romance Movie
    </button>
  </div>

  <div id="roulette-card" style="background: #181818; border-radius: 12px; padding: 22px; border-left: 6px solid #FF2A54; max-width: 680px; margin: 0 auto; text-align: left; display: none; box-shadow: 0 6px 20px rgba(0,0,0,0.5);">
    <div style="display: flex; justify-content: space-between; align-items: baseline; flex-wrap: wrap; gap: 10px; margin-bottom: 12px;">
      <h3 id="r-title" style="color: #ffffff; margin: 0; font-size: 1.4rem;">Movie Title</h3>
      <div style="display: flex; gap: 8px; flex-wrap: wrap;">
        <span id="r-genre" style="background: #FF2A54; color: #fff; padding: 3px 10px; border-radius: 6px; font-size: 0.8rem; font-weight: bold;">Action</span>
        <span id="r-year" style="background: #2a2a2a; color: #fff; padding: 3px 10px; border-radius: 6px; font-size: 0.8rem;">2020</span>
        <span id="r-rating" style="background: #2a2a2a; color: #bbb; padding: 3px 10px; border-radius: 6px; font-size: 0.8rem;">PG-13</span>
        <span id="r-duration" style="background: #2a2a2a; color: #bbb; padding: 3px 10px; border-radius: 6px; font-size: 0.8rem;">115 min</span>
      </div>
    </div>
    <p id="r-desc" style="color: #d1d5db; margin: 0; line-height: 1.6; font-size: 1rem;">Movie synopsis will appear here.</p>
  </div>
</div>

<script>
function spinMovie(genre) {
  if (typeof filmsData === 'undefined') return;
  const pool = filmsData.filter(m => m.genre === genre);
  if (!pool.length) return;
  const picked = pool[Math.floor(Math.random() * pool.length)];

  const card = document.getElementById('roulette-card');
  const title = document.getElementById('r-title');
  const genreBadge = document.getElementById('r-genre');
  const year = document.getElementById('r-year');
  const rating = document.getElementById('r-rating');
  const duration = document.getElementById('r-duration');
  const desc = document.getElementById('r-desc');

  title.textContent = picked.title;
  genreBadge.textContent = picked.genre;
  year.textContent = picked.release_year;
  rating.textContent = picked.rating || 'Unrated';
  duration.textContent = picked.duration || '';
  desc.textContent = picked.description;

  const color = picked.genre === 'Action' ? '#FF2A54' : '#00E5FF';
  const textCol = picked.genre === 'Action' ? '#FFFFFF' : '#0a0a0a';
  card.style.borderLeftColor = color;
  genreBadge.style.background = color;
  genreBadge.style.color = textCol;

  card.style.display = 'block';
}

document.getElementById('btn-action').addEventListener('click', () => spinMovie('Action'));
document.getElementById('btn-romance').addEventListener('click', () => spinMovie('Romance'));
</script>
```

## Movie Explorer: Search & Filter

Want to see which movies are behind the numbers? Use the search bar below to find your favorite films, or filter by genre and year:

```{r}
#| label: movie-table
films |>
  select(Title = title, Genre = genre, Year = release_year, Rating = rating, Duration = duration, Description = description) |>
  datatable(
    filter = "top",
    options = list(
      pageLength = 8,
      autoWidth = TRUE,
      searchHighlight = TRUE
    ),
    rownames = FALSE,
    class = "cell-border stripe"
  )
```


